---
title: "STAT 331 Portfolio"
author: "Arturo Ordaz-Gutierrez"
theme: pulse
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1

```{r}
#| label: wd-1-csv-1
# Lab 2: Question 1

surveys <- read_csv(here::here("Week2", "Lab2", "surveys.csv"))
```

-   `csv` Example 2

```{r}
#| label: wd-1-csv-2
# Lab 3: Question 2

teacher_evals <- read_csv(here::here('Week3', 'Lab3', 'teacher_evals.csv'))
```

-   `xlsx`

```{r}
#| label: wd-1-xlsx
# Week 2: Check-in 2.2

ages <- read_xlsx(here::here("Week2",
                             "Check-in",
                             "Ages_Data",
                             "ages.xlsx"),
                  sheet = "ages", 
                  skip  = 0, 
                  n_max = 6)
```

-   `txt`

```{r}
#| label: wd-1-txt
# Week 2: Check-in 2.3

ages_space <- read_table(here::here("Week2",
                                    "Check-in",
                                    "Ages_Data",
                                    "ages_space.txt"),
                         col_names = TRUE)
```


**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns

```{r}
#| label: wd-2-ex-1
# Lab 3: Question 5

teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    teacher_id   = as.character(teacher_id),
    course_id    = as.character(course_id)
  ) |>
  select(
    course_id, teacher_id, question_no, no_participants, resp_share,
    SET_score_avg, percent_failed_cur, academic_degree, seniority, sex
  )
```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2

```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`)

```{r}
#| label: wd-2-ex-3

```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-3-numeric-ex-1
# Lab 3: Question 5

teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    teacher_id   = as.character(teacher_id),
    course_id    = as.character(course_id)
  ) |>
  select(
    course_id, teacher_id, question_no, no_participants, resp_share,
    SET_score_avg, percent_failed_cur, academic_degree, seniority, sex
  )
```

-   Numeric Example 2

```{r}
#| label: wd-3-numeric-ex-1
# Lab 4: Question 5

ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(census_region, study_year) |>
  summarise(median_income_2018 = median(mhi_2018, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = study_year, 
              values_from = median_income_2018, 
              names_prefix = "Year: ") |>
  arrange(desc('Year: 2018'))
```

-   Character Example 1 (any context)

```{r}
#| label: wd-3-character
# Lab 4: Question 2

ca_childcare <- counties |>
  filter(state_name == "California") |>
  inner_join(childcare_costs,
             by = join_by(county_fips_code == county_fips_code))
```

-   Character Example 2 (example must use functions from **stringr**)

```{r}
#| label: wd-3-string
# Lab 5: Identify the Two Witnesses 

# Witness 1: Last house on Northwestern Dr
witness1 <- person |>
  filter(address_street_name == "Northwestern Dr") |>
  arrange(desc(address_number)) |>
  head(1)

# Witness 2: Annabel on Franklin Ave
witness2 <- person |>
  filter(str_detect(name, "Annabel"),
         address_street_name == "Franklin Ave")

witness1
witness2
```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date
# Lab 5: Gym Activity

get_fit_now_check_in |>
  mutate(check_in_date = ymd(check_in_date)) |>
  filter(membership_id %in% suspect_member$id,
         check_in_date == ymd("2018-01-09"))
```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-4-numeric-ex-1
# Lab 3: Question 7

teacher_evals_clean |>
  distinct(teacher_id, .keep_all = TRUE) |>
  count(academic_degree, sex) |>
  mutate(percent = round(100 * n / sum(n), 1)) |>
  arrange(desc(percent))
```

-   Numeric Example 2

```{r}
#| label: wd-4-numeric-ex-2
# Lab 5: Find Who Hired Jeremy Bowers

concert_attendees <- facebook_event_checkin |>
  filter(event_name == "SQL Symphony Concert",
         str_starts(date, "201712")) |>
  count(person_id, name = "n_checkins") |>
  filter(n_checkins >= 3)

concert_attendees
```

-   Factor Example 1 (renaming levels)

```{r}
#| label: wd-4-factor-ex-1
# Lab 4: Question 4

ca_childcare <- ca_childcare |> 
  mutate(
    county_name = str_remove(county_name, " County"),
    census_region = fct_collapse(
      .f = county_name,
      superior      = superior_counties,
      north_coast   = north_coast_counties,
      san_francisco = san_fran_counties,
      n_san_joaquin = n_san_joaquin_counties,
      central_coast = central_coast_counties,
      s_san_joaquin = s_san_joaquin_counties,
      inland        = inland_counties,
      la            = la_county,
      orange        = orange_county,
      san_diego_imperial = san_diego_imperial_counties
    )
  )
```

-   Factor Example 2 (reordering levels)

```{r}
#| label: wd-4-factor-ex-2
# Lab 5: Find Initial Suspect

suspect_member <- get_fit_now_member |>
  mutate(
    # Convert membership_status to an ordered factor
    membership_status = factor(
      membership_status,
      levels = c("bronze", "silver", "gold"),
      ordered = TRUE
    )
  ) |>
  filter(str_starts(id, "48Z"),
         membership_status == "gold")
```

-   Character (example must use functions from **stringr**)

```{r}
#| label: wd-4-string
# Lab 5: Identify the Two Witnesses

filter(str_detect(name, "Annabel"),
       address_street_name == "Franklin Ave")
```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date
# lab 5: Find Initial Suspect

suspect_checkin <- get_fit_now_check_in |>
  mutate(check_in_date = ymd(check_in_date)) |>
  filter(membership_id %in% suspect_member$id,
         check_in_date == ymd("2018-01-09"))

suspect_checkin
```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1

```{r}
#| label: wd-5-left-ex-1
# Lab 4: Question 3

ca_childcare <- ca_childcare |>
  left_join(tax_rev,
            by = join_by(county_name == entity_name,
                         study_year == year))
```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

```

-   `left_join()` **or** `right_join()` Example 2

```{r}
#| label: wd-5-left-right-ex-2
#Lab 5: Joining to Find Specific Person

suspect_pool <- concert_attendees |>
  left_join(person, by = c("person_id" = "id")) |>
  left_join(drivers_license, by = c("license_id" = "id"))
```

-   `inner_join()` Example 1

```{r}
#| label: wd-5-inner-ex-1
# Lab 4: Question 2

ca_childcare <- counties |>
  filter(state_name == "California") |>
  inner_join(childcare_costs,
             by = join_by(county_fips_code == county_fips_code))
```

-   `inner_join()` Example 2

```{r}
#| label: wd-5-inner-ex-2
# Lab 5: Interview Initial Suspect

suspect_interview <- suspect |>
  inner_join(interview, by = c("person_id" = "person_id"))
```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long
# Lab 3: Question 8
teacher_evals_clean |>
  filter(if_any(everything(), is.na)) |>
  summarize(across(everything(), ~ sum(is.na(.))), .groups = "drop") |>
  pivot_longer(everything(), names_to = "variable", values_to = "num_missing") |>
  filter(num_missing > 0)
```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide
# Lab 4: Question 5

income_2008_2018 <- ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(census_region, study_year) |>
  summarise(
    median_income_2018_dollars = median(mhi_2018, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = study_year,
    values_from = median_income_2018_dollars,
    names_prefix = "Median income (2018 dollars) - "
  ) |>
  arrange(desc(`Median income (2018 dollars) - 2018`))
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Example 1
-   Example 2
-   Example 3
-   Example 4
-   Example 5

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1
# Lab 4: Question 7

library(RColorBrewer)

recreate_plot <- ca_childcare |> 
  select(
    county_name, 
    study_year, 
    mc_infant, 
    mc_toddler, 
    mc_preschool, 
    census_region
  ) |>
  pivot_longer(
    cols = mc_infant:mc_preschool,
    names_to = "age",
    values_to = "price"
  ) |> 
  mutate(
    age = case_when(
      age == "mc_infant" ~ "Infant",
      age == "mc_toddler" ~ "Toddler",
      age == "mc_preschool" ~ "Preschool",
      TRUE ~ age
    ),
    census_region = fct_reorder2(
      .f = census_region, 
      .x = study_year, 
      .y = price, 
      .desc = TRUE
    )
  )

mycolors <- colorRampPalette(brewer.pal(7, "Accent"))(10)

ggplot(
  data = recreate_plot,
  mapping = aes(x = study_year, y = price, color = census_region)
) +
  geom_point(size = 0.75, alpha = 0.6) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  facet_wrap(~factor(age, levels = c("Infant", "Toddler", "Preschool"))) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", big.mark = ",")) +
  scale_x_continuous(
    breaks = seq(2008, 2018, by = 2),
    limits = c(2008, 2018)
  ) +
  scale_color_manual(values = mycolors) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare ($)",
    x = "Study Year",
    y = NULL,
    color = "California Region"
  ) +
  theme_bw(base_size = 10) +
  theme(
    aspect.ratio = 1.25,
    axis.text.x = element_text(angle = 30)
  )
```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2
# Lab 3: Question 5

teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    teacher_id = as.character(teacher_id),
    course_id  = as.character(course_id)
  ) |>
  select(
    course_id,
    teacher_id,
    question_no,
    no_participants,
    resp_share,
    SET_score_avg,
    percent_failed_cur,
    academic_degree,
    seniority,
    sex
  )
```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example
# Lab 3: Question 5

#| label: data-cleaning
teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    teacher_id = as.character(teacher_id),
    course_id = as.character(course_id)
  ) |>
  select(
    course_id, 
    teacher_id, 
    question_no, 
    no_participants, 
    resp_share,
    SET_score_avg, 
    percent_failed_cur, 
    academic_degree, 
    seniority, 
    sex
  )
```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num
# Lab 4: Question 8

ggplot(
  data = ca_childcare,
  mapping = aes(x = mhi_2018, y = mc_infant)
) +
  geom_point(
    alpha = 0.5,
    color = "darkgreen"
  ) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "gold"
  ) +
  labs(
    x = "Median household income (2018 dollars)",
    y = "Median weekly price (infant, center-based)",
    title = "Income vs. Infant Center-Based Weekly Price (CA, 2018)"
  ) +
  theme_bw(base_size = 10)
```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat
# lab 4: Question 7

ggplot(
  data = recreate_plot,
  mapping = aes(x = study_year, y = price, color = census_region)
) +
  geom_point(
    size = 0.75,
    alpha = 0.6
  ) +
  geom_smooth(
    linewidth = 0.75,
    se = TRUE
  ) +
  facet_wrap(
    ~factor(age, c("Infant", "Toddler", "Preschool"))
  ) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "$", big.mark = ",")
  ) +
  scale_x_continuous(
    breaks = seq(2008, 2018, by = 2),
    limits = c(2008, 2018)
  ) +
  scale_color_manual(
    values = mycolors
  ) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare ($)",
    x = "Study Year",
    y = NULL,
    color = "California Region"
  ) +
  theme_bw(base_size = 10) +
  theme(
    aspect.ratio = 1.25,
    axis.text.x = element_text(angle = 30)
  )
```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat
# Lab 3: Question 7

ggplot(
  data = teacher_evals_clean |>
    distinct(teacher_id, .keep_all = TRUE) |>
    count(academic_degree, sex),
  mapping = aes(x = academic_degree, y = n, fill = sex)
) +
  geom_col(
    position = "dodge"
  ) +
  labs(
    title = "Count of Instructors by Degree and Sex",
    x = "Academic Degree",
    y = "Number of Instructors",
    fill = "Sex"
  ) +
  theme_bw(base_size = 10)
```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date
# lab 4: Question 7

ggplot(
  data = recreate_plot,
  mapping = aes(x = study_year, y = price, color = census_region)
) +
  geom_line() +
  geom_smooth(
    se = TRUE,
    alpha = 0.2
  ) +
  scale_x_continuous(
    breaks = seq(2008, 2018, by = 2)
  ) +
  labs(
    x = "Study Year",
    y = "Median Weekly Price ($)",
    title = "Change in Childcare Costs Over Time by Region",
    color = "Region"
  ) +
  theme_bw(base_size = 10)
```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1
# Lab 4: Question 7

ggplot(
  data = recreate_plot,
  mapping = aes(x = study_year, y = price, color = census_region)
) +
  geom_point(
    size = 0.75,
    alpha = 0.6
  ) +
  geom_smooth(
    linewidth = 0.75,
    se = TRUE
  ) +
  facet_wrap(
    ~factor(age, c("Infant", "Toddler", "Preschool"))
  ) +
  scale_y_continuous(
    labels = scales::dollar_format(prefix = "$", big.mark = ",")
  ) +
  scale_x_continuous(
    breaks = seq(2008, 2018, by = 2),
    limits = c(2008, 2018)
  ) +
  scale_color_manual(
    values = mycolors
  ) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare ($)",
    x = "Study Year",
    y = NULL,
    color = "California Region"
  ) +
  theme_bw(base_size = 10) +
  theme(
    aspect.ratio = 1.25,
    axis.text.x = element_text(angle = 30)
  )
```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2
# Lab 4: Question 7

library(RColorBrewer)
mycolors <- colorRampPalette(brewer.pal(7, "Accent"))(10)
```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3
# Lab 4: Question 8

ggplot(
  data = ca_childcare,
  mapping = aes(x = mhi_2018, y = mc_infant)
) +
  geom_point(
    alpha = 0.5,
    color = "darkgreen"
  ) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "gold"
  ) +
  labs(
    title = "Income vs. Infant Center-Based Weekly Price (CA, 2018)",
    x = "Median household income (2018 dollars)",
    y = "Median weekly price (infant, center-based)"
  ) +
  theme_bw(base_size = 10)
```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4
# Lab 4: Question 7

theme(
  aspect.ratio = 1.25,
  axis.text.x = element_text(angle = 30)
)
```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5
# Lab 4: Question 7

recreate_plot <- ca_childcare |>
  select(county_name, study_year, mc_infant, mc_toddler, mc_preschool, census_region) |>
  pivot_longer(
    cols = mc_infant:mc_preschool,
    names_to = "age",
    values_to = "price"
  ) |>
  mutate(
    age = case_when(
      age == "mc_infant" ~ "Infant",
      age == "mc_toddler" ~ "Toddler",
      age == "mc_preschool" ~ "Preschool",
      TRUE ~ age
    ),
    census_region = fct_reorder2(
      .f = census_region,
      .x = study_year,
      .y = price,
      .desc = TRUE
    )
  )
```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1

```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2

```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize
# Lab 3: Question 7

teacher_evals_clean |>
  distinct(teacher_id, .keep_all = TRUE) |>
  summarize(
    total_instructors = n(),
    avg_seniority = mean(seniority, na.rm = TRUE)
  )
```

-   Example using `across()`

```{r}
#| label: dvs-4-across
# Lab 3: Question 8

teacher_evals_clean |>
  filter(if_any(everything(), is.na)) |>
  summarize(
    across(everything(), ~ sum(is.na(.))),
    .groups = "drop"
  )

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

```

-   Example 2

```{r}
#| label: dvs-5-2

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1

```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2

```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3

```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`)

```{r}
#| label: dvs-7-ex-2

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

```

-   using `across()`

```{r}
#| label: pe-1-across

```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1

```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2

```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

```

-   Example 2

```{r}
#| label: dsm-1-2

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

```

-   Example 2

```{r}
#| label: dsm-2-2

```

-   Example 3

```{r}
#| label: dsm-2-3

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->
